# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
Assessment data will be loaded directly from zip file:
```{r}
assessment.data <- read.csv(unzip("activity.zip", "activity.csv"))
```
Lets examine structure of crated data frame:
```{r}
str(assessment.data)
```
We can see that data type for **date** is *Factor*, and it probably should be *Date*. Let's adjust that:
```{r}
assessment.data$date <- as.Date(assessment.data$date)
```
Simple summary of readed data:
```{r}
summary(assessment.data)
```
## What is mean total number of steps taken per day?
Sum of steps for each day. Steps variable is **x**:
```{r}
total.steps.day <- aggregate.data.frame(x = assessment.data$steps, by = list(Date = assessment.data$date), sum)
total.steps.day
```
Histogram:
```{r}
hist(total.steps.day$x)
```

Mean and median values:
```{r}
mean.steps <- mean(total.steps.day$x, na.rm = T)
median.steps <- median(total.steps.day$x, na.rm = T)
mean.steps
median.steps
```

## What is the average daily activity pattern?
Lets load plotting library:
```{r}
library("ggplot2")
```
Sum of steps for each interval. Steps variable is **x**:
```{r}
total.steps.interval <- aggregate.data.frame(x = assessment.data$steps, by = list(Interval = assessment.data$interval), sum)
total.steps.interval
```
Plot:
```{r}
ggplot()
```
## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
