# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
Assessment data will be loaded directly from zip file:
```{r}
assessment.data <- read.csv(unzip("activity.zip", "activity.csv"))
```
Lets examine structure of crated data frame:
```{r}
str(assessment.data)
```
We can see that data type for **date** is *Factor*, and it probably should be *Date*. Let's adjust that:
```{r}
assessment.data$date <- as.Date(assessment.data$date)
```
Lets also convert interval to timestamp format. First, lets use strptime to extract timestamp:
```{r}
assessment.data$interval <- as.integer(as.POSIXct(strptime(
  formatC(assessment.data$interval, width = 4, format = "d", flag = "0"), 
  format = "%H%M")))
```
And then subtract first timestamp to have proper timing:
```{r}
assessment.data$interval <- assessment.data$interval - assessment.data$interval[1]
```
Simple summary of read data:
```{r}
summary(assessment.data)
```
## What is mean total number of steps taken per day?
Sum of steps for each day. Steps variable is **x**:
```{r}
total.steps.day <- aggregate.data.frame(x = assessment.data$steps, by = list(Date = assessment.data$date), sum, na.rm = TRUE)
head(total.steps.day)
```
Histogram:
```{r}
hist(total.steps.day$x, main = "Histogram of total steps per day", xlab = "Sum of steps per day")
```

Mean and median values:
```{r}
mean.steps <- mean(total.steps.day$x, na.rm = T)
median.steps <- median(total.steps.day$x, na.rm = T)
mean.steps
median.steps
```

## What is the average daily activity pattern?
Lets load plotting library:
```{r}
library("ggplot2")
```
Average of steps for each interval. Steps variable is **x**:
```{r}
mean.steps.interval <- aggregate.data.frame(x = assessment.data$steps, by = list(Interval = assessment.data$interval), mean, na.rm = TRUE)
head(mean.steps.interval)
```
Plot:
```{r}
ggplot(mean.steps.interval, aes(Interval/3600,x)) + geom_line() 
```

Time interval which has highest average steps count:
```{r}
intense.time <- mean.steps.interval$Interval[which.max(mean.steps.interval$x)]
format(as.POSIXct('0001-01-01 00:00:00') + intense.time, "%I:%M:%S %p") 
```
## Imputing missing values
### Number of missing values:
We know from previous summary, that column which can have NA's is column *steps*. So to know number of missing cases, we only need to examine NA's in *steps*:
```{r}
total.rows <- nrow(assessment.data)
missing.rows <- sum(is.na(assessment.data$steps))
percent.missing <- (missing.rows/total.rows)*100
```
Results are: we have `r total.rows` total rows, from which `r missing.rows` contains missing values - that is `r percent.missing` percent of all rows.
### Filling missing values:
I decided to fill missing values based on mean value for that time interval. For each missing value, its interval will be taken. For this interval, number of mean steps *x* will be taken from calculated previously *mean.steps.interval* data frame.
### Creating new dataset with filled missing values
```{r}
new.data <- assessment.data
for (i in which(is.na(new.data$steps))) {
  new.data$steps[i] <- mean.steps.interval$x[mean.steps.interval$Interval == new.data$interval[i]]
}
head(new.data)
```
### Comparing with data containing NA
I will run the same analytics like in second paragraph:
Sum of steps for each day. Steps variable is **x**:
```{r}
total.steps.day <- aggregate.data.frame(x = new.data$steps, by = list(Date = new.data$date), sum)
head(total.steps.day)
```
Histogram:
```{r}
hist(total.steps.day$x, main = "Histogram of total steps per day", xlab = "Sum of steps per day")
```

Mean and median values:
```{r}
mean.steps <- mean(total.steps.day$x)
median.steps <- median(total.steps.day$x)
mean.steps
median.steps
```
After filling missing values, both mean and median are slightly higher.
## Are there differences in activity patterns between weekdays and weekends?
### Adding new factor variable to dataset:
```{r}
Sys.setlocale(category = "LC_TIME", locale = "C") # fixing locales
new.data$day <- weekdays(new.data$date)
new.data$day <- ifelse(new.data$day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), "weekday", "weekend")

```
### Calculating separate means for weekend and weekdays:
Aggregation method as previously. I will create two separated data frames with means for each day type and them merge them togeather:
```{r}
mean.steps.interval.weekend <- aggregate.data.frame(x = new.data$steps[new.data$day == "weekend"], by = list(Interval = new.data$interval[new.data$day == "weekend"]), mean)
head(mean.steps.interval.weekend)

mean.steps.interval.weekday <- aggregate.data.frame(x = new.data$steps[new.data$day == "weekday"], by = list(Interval = new.data$interval[new.data$day == "weekday"]), mean)
head(mean.steps.interval.weekday)
merged.mean.steps <- merge(mean.steps.interval.weekday, mean.steps.interval.weekend, by = "Interval")
names(merged.mean.steps) <- c("interval", "weekday", "weekend")
```
Plotting:
```{r}
par(mfrow=c(2,1))
plot(merged.mean.steps$interval/3600,merged.mean.steps$weekday, main="Mean steps for each interval during weekdays", type = "l", xlab = "Interval", ylab = "Mean number of steps")
plot(merged.mean.steps$interval/3600,merged.mean.steps$weekend, main="Mean steps for each interval during weekends", type = "l", xlab = "Interval", ylab = "Mean number of steps")
```